---
title: "Task 1 – Exploring Data Types in R"
format: html
---

## 1.1 Atomic Types

Let's define attributes of our fictional planet **Aqualis**.

```{r}
# Numeric values
temperature <- 24.5
gravity <- 9.8
ocean_depth <- 3500.2
radiation_level <- 0.87

# Integer values
number_of_moons <- 2L
cities_count <- 8L
research_stations <- 3L

# Logical values
has_water <- TRUE
is_inhabited <- TRUE

# Character values
planet_name <- "Aqualis"
region <- "Hydros Sector"
category <- "Ocean Planet"

# Describe them in a sentence
cat(
  planet_name, "is located in", region, 
  "and is classified as an", category,
  "with", number_of_moons, "moons and average temperature", temperature, "°C.\n"
)
```

---

## 1.2 Vectors

```{r}
# Numeric + integer vector
numeric_vector <- c(temperature, gravity, ocean_depth, radiation_level, 
                    number_of_moons, cities_count, research_stations)

# Sequence vector
sequence_vector <- seq(20, 100, by = 10)

# Logical vector (10 random TRUE/FALSE)
set.seed(42)
logical_vector <- sample(c(TRUE, FALSE), 10, replace = TRUE)

numeric_vector
sequence_vector
logical_vector
```

---

## 1.3 Matrix

```{r}
# Create 3x3 matrix
set.seed(123)
planet_matrix <- matrix(sample(-50:50, 9, replace = TRUE), nrow = 3, ncol = 3)

# Name rows and columns
rownames(planet_matrix) <- c("Ocean", "Land", "Atmosphere")
colnames(planet_matrix) <- c("Pressure", "Temperature", "Density")

planet_matrix

# Row and column means
rowMeans(planet_matrix)
colMeans(planet_matrix)
```

---

## 1.4 List

```{r}
planet_list <- list(
  temperature = temperature,
  gravity = gravity,
  ocean_depth = ocean_depth,
  numeric_vector = numeric_vector,
  sequence_vector = sequence_vector,
  logical_vector = logical_vector,
  matrix = planet_matrix,
  description = "Aqualis is an ocean-covered exoplanet with moderate gravity and thriving research stations."
)

planet_list
```

---

## 1.5 Factor

```{r}
habitat_levels <- c("poor", "average", "rich", "average", "poor", "rich")
habitat_factor <- factor(habitat_levels, levels = c("poor", "average", "rich"))
table(habitat_factor)
```

---

## 1.6 Data Frame

```{r}
planet_df <- data.frame(
  region = c("North", "South", "East", "West", "Central"),
  population_mil = c(3.4, 2.1, 1.8, 4.2, 2.9),
  is_habitable = c(TRUE, TRUE, TRUE, FALSE, TRUE),
  resource_level = c("high", "medium", "medium", "low", "high")
)

# Rename columns
colnames(planet_df) <- c("Region", "Population_mil", "Habitable", "Resource_Level")

# Structure and summary
str(planet_df)
summary(planet_df)

# Derived column (resources per person)
planet_df$Resource_Index <- ifelse(planet_df$Resource_Level == "high", 1.5,
                            ifelse(planet_df$Resource_Level == "medium", 1.0, 0.5))

planet_df

